Modelling last glacial cycle ice dynamics in the Alps

### 2018-07-31 articles/cycle/response-editor.tex

% response-editor.tex
% ----------------------------------------------------------------------
\input{response-header}

\title{Authors' response to the Editor}
\author{J. Seguinot, on behalf of all authors.}
%\date{}

\begin{document}
\maketitle
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

    Dear Andreas Vieli,

    We apologize for delays accumulated during the peer review of our
    manuscript. We believe that we have addressed all points raised by the
    reviews and hereby submit our revised version to \emph{The Cryosphere}.
    Please find hereafter a short summary and a marked-up file listing
    changes made to the manuscript since its first publication in \emph{The
    Cryosphere Discussion}. Please refer to our public reponses for more
    detail explanations.

    \begin{itemize}

        \item \textbf{Sect.~2 (Ice sheet model set-up):}
        We have made explicit the lack of measurements on ice softness for
        water contents above 1\,\%, clarified that subglacial water is not
        transported, explained the effect of erroneous elastic thickness, and
        justified our value for the air temperature lapse-rate.

        \item \textbf{Sect.~3 (Palaeo-climate forcing):}
        We have provided a short review on available palaeo-climate proxy
        records in or near the Alps, summarized the state of knowledge on past
        precipitation changes, and pointed out inconsistencies between our
        reference climate forcing and other data.

        \item \textbf{Sect.~4 (Results and discussion):}
        We have pointed out inconsistent LGM ice extent reconstructions in the
        south-western Alps, removed the discussion on self-sustained ice domes,
        and discussed uncertainties regarding our conclusion on trimlines.
        \emph{New references were added following private comments from
              Christopher~Carcaillet and Wilfried~Haeberli}.

        \item \textbf{Sect.~5 (Conclusions):}
        We have added a word of caution regarding our conclusion on EPICA being
        the optimal forcing, listed major sources of uncertainties, and
        emphasized the need for more realistic climate forcing in future
        studies.

        \item \textbf{Fig.~1:}
        We have removed self-sustained ice domes and updated our somewhat
        arbitrary choice of transfluence locations.

        \item \textbf{Fig.~6:}
        We have corrected histogram data and replaced areas modelled to have
        experienced temperate ice for less than 1\,ka by frozen-based areas
        at the time of maximum ice thicknes.

        \item \textbf{Various:}
        \emph{We found that thicker 200\,m contour lines were disturbing other
              figure elements and eventually decided to keep original thickness
              inconsistently with our response to referee~\#2.}
        We have implemented numerous suggestions for text improvements.

    \end{itemize}

    We thank you very much again for your editorial work for our manuscript.

\end{document}

### 2018-07-31 articles/cycle/alps-cycle.tex
Add refs on Alpine tree refugia.
Add reponse to the editor.

    In the upper Rhone Valley, the maximum ice surface elevation reached
    during MIS~2 in our simulation, corrected for bedrock deformation, is
    consistently modelled to have occured several hundred metres above the
    observed trimline elevations (Fig.~\ref{fig:trimlines}a and b), with a mean
    difference of 861\,m (and a standard deviation 197\,m,
    Fig.~\ref{fig:trimlines}b), which is significant in respect to the modelled
    maximum ice thickness in the Rhone Valley (and the Alps) of 2586\,m. This
    result depends on uncertain basal sliding and ice rheological parameters
    to which the model sensitivity was not tested here. Unfortunately,
    validation through the bedrock uplift rate \citep[cf.][]{Kuchar.etal.2012}
    is not doable in the Alps due its lower values, active tectonics and
    uncertainties on geological properties \citep[cf.][]{Mey.etal.2016}.
    However, regional sensitivity tests show that a modelled surface compatible
    with the trimline elevations is incompatible with the documented
    overspilling flow across the Simplon Pass \citep{Becker.etal.2017}, and LGM
    palaeo-climate proxy records \citep{Cohen.etal.2017}.

@@

    We are very thankful to Constantine Khroulev, Ed Bueler, and Andy
    Aschwanden for providing constant help and development with PISM, in
    particular with recent issues with the computation of ice temperature
    (Github issue no.~371) and with the computation of bedrock deformation
    (Github issues no.~370 and 377). We are equally thankful to three
    anonymous reviewers and Giovanni Monagato for very constructive inputs
    during peer-review, and to Marc Luetscher for insightful discussions of
    preliminary results during the EGU~General Assembly~2017. The experimental
    design and article layout used here are based on previous work which
    received much guidance from Irina Rogozhina and Arjen~P. Stroeven.
    The current work was supported by the Swiss National Science Foundation
    grants no.~200020-169558 and 200021-153179/1 to M.~Funk. Computer resources
    were provided by the Swiss National Supercomputing Centre (CSCS)
    allocations no.~s573 and sm13 to J.~Seguinot.

>>>>

    In the upper Rhone Valley, the maximum ice surface elevation reached
    during MIS~2 in our simulation, corrected for bedrock deformation, is
    consistently modelled to have occured several hundred metres above the
    observed trimline elevations (Fig.~\ref{fig:trimlines}a and b), with a mean
    difference of 861\,m (and a standard deviation 197\,m,
    Fig.~\ref{fig:trimlines}b), which is significant in respect to the modelled
    maximum ice thickness in the Rhone Valley (and the Alps) of 2586\,m. This
    result depends on uncertain basal sliding and ice rheological parameters
    to which the model sensitivity was not tested here. Unfortunately,
    validation through the bedrock uplift rate \citep[cf.][]{Kuchar.etal.2012}
    is not doable in the Alps due its lower values, active tectonics and
    uncertainties on geological properties \citep[cf.][]{Mey.etal.2016}.
    Interestingly, tree and fire remains evidence nunataks in the south-western
    Alps by 21\,ka \citep{Carcaillet.Blarquez.2017, Carcaillet.etal.2018}.
    However, regional sensitivity tests show that a modelled surface compatible
    with the trimline elevations is incompatible with the documented
    overspilling flow across the Simplon Pass \citep{Becker.etal.2017}, and LGM
    palaeo-climate proxy records \citep{Cohen.etal.2017}.

@@

    We are very thankful to Constantine Khroulev, Ed Bueler, and Andy
    Aschwanden for providing constant help and development with PISM, in
    particular with recent issues with the computation of ice temperature
    (Github issue no.~371) and with the computation of bedrock deformation
    (Github issues no.~370 and 377). We are equally thankful to three
    anonymous reviewers, Giovanni Monagato, and the editor Andreas Vieli for
    their careful read and constructive inputs during peer-review.
    We thank Marc Luetscher for an insightful discussion of
    preliminary results during the EGU~General Assembly~2017. The experimental
    design and article layout used here are based on previous work which
    received much guidance from Irina Rogozhina and Arjen~P. Stroeven.
    The current work was supported by the Swiss National Science Foundation
    grants no.~200020-169558 and 200021-153179/1 to M.~Funk. Computer resources
    were provided by the Swiss National Supercomputing Centre (CSCS)
    allocations no.~s573 and sm13 to J.~Seguinot.
